---
title: "BI586_Rlab_Worksheet_mingers"
output: "word_document"
---
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

- Load the data file you will be using for the analysis and give it a name so as to access it more easily
- Also, change the name of column 21 to "Average WD"
```{r}
Data=read.csv(file = "Wasting_Data.csv",skip = 1, stringsAsFactors = T)
colnames(Data)[21] = "Average.WD"
```


Problem 9: Write code that will print the column name for column 3 and the rowname for row 4
```{r}
colnames(Data)[3]
rownames(Data)[4]
col3 = colnames(Data)[3]
row4 = rownames(Data)[4]
col3_row4 = c(col3 , row4)
col3_row4
```


- Sometimes you will want to subset a dataframe to include only specific columns or only certain values 
- In our wasting data file we have a column called Transect. 
- We can view the values of this column by simply clicking on the variable "Data" in our global environment OR we can have R print out all the values of this column
- If you don't want R to barf up all the values but you are interested in how many unique values there are for this column, you can look at this column's levels

Problem 10: How many levels does the column "Transect" have?
```{r}
levels(Data$Transect)
```
- This function shows us that there are four levels in the column transect

- Lets say we want to create a dataframe that only has data for transect A, we can do this a few different ways
- One way is to use the filter function from the package dplyr
```{r}
TransectA=filter(Data, Transect == "A")
head(TransectA)
```


Problem 11: How many rows are in your new dataframe TransectA?
```{r}
nrow(TransectA)
```
- There are 10 rows in TransectA


Problem 12: Create a dataframe that consistes of only data from Transect A AND Transect B. How many rows are in this new dataframe?
```{r}
TransectAB = filter(Data, Transect == "A" | Transect =="B")
nrow(TransectAB)
```
- There are 20 rows in TransectAB

- Maybe we don't care about most of these columns and want to just work with the first three
```{r}
InfoTransectA=TransectA[,1:3]
head(InfoTransectA)
```


How do we save stuff that we can create in R?

- Even though you can save your script with a simple ctrl+S, this is just saving the instructions
- Sometimes we want to save an output so we don't have to run through our whole code again to look at it.
- One way of doing this is saving EVERYTHING in your global environment by using the following:
```{r}
save.image(file="filename.RData")
```

- Then the function load can be used to re-load your saved global environment
```{r}
load(file="filename.RData")
```

- You can also save files as .csv or a space separated file using the following functions:
```{r}
write.csv(InfoTransectA, file="InfoTranectA.csv")
write.table(InfoTransectA, file="InfoTransectA.txt",sep=" ")
write.table(InfoTransectA, file="InfoTransectA.tab", sep="\t")
```


